Index measures for oak decline severity using phenotypic descriptors

### 2019-12-14 manuscript/manuscript.Rmd
group decline index section finished

```{r loadData, include = FALSE}
loadd(site_rf_margin,
      site_rf_post_correction_margin,
      PDI_status_ttest,
      DAI_groups_ttest,
      PDI_rf_tune_params,
      DAI_rf_tune_params,
      analysis_suitable_data,
      site_corrected_analysis_suitable_data)
```

@@

Descriptors were collected relating to the trees shape, size, crown condition and the presence of biotic agents including insects, bacteria and fungi.
All trees were manually assigned a decline status category; control, AOD, COD or remission prior to collection of phenotypic descriptors.
Diameter at breast height (DBH) was collected at a stem height of 1.3m, to the nearest 0.1cm.
Total height, timber hight and lowest live crown were measured using an inclinometer to the nearest 0.1m
The timber height was measured at either 7cm top stem diameter or where no main stem could be distinguished.
For the crown, canopy closure, missing crown and absolute crown density percentages were visually estimated within 5% classes

@@

### Oak decline indexes

Trees with a score around 0 are neither distinctively AOD or COD sharing characteristics of both conditions in trees with PDI greater than 0.3.
A significant difference was found in the PDI between non-symptomatic (control) and symptomatic (AOD and COD) decline statuses using an independent two sample t-test (`r glue("t({PDI_status_ttest$parameter}) = {signif(PDI_status_ttest$statistic,digits = 3) %>% abs()}, p < 0.001")`) with group means of `r signif(PDI_status_ttest$estimate1,digits = 3)` and `r signif(PDI_status_ttest$estimate2,digits = 3)` for non-symptomatic and symptomatic statuses respectively.
Significance was also found in the DAI between the AOD and COD groups (`r glue("t({DAI_groups_ttest$parameter}) = {signif(DAI_groups_ttest$statistic,digits = 3) %>% abs()}, p < 0.001")`) with group means of `r signif(DAI_groups_ttest$estimate1,digits = 3)` and `r signif(DAI_groups_ttest$estimate2,digits = 3)` for the AOD and COD groups respectively.
A scatter plot of PDI against DAI is shown in Figure # for all 174 trees.
It can clearly be seen that the

### Descriptor contributions and trends to the decline indexes

- Trends of key phenotypic descriptors withing the PDI and DAI measures
- Agrilus exit hole density had no contribution
- Lime results
- Top ranked descriptor index response surfaces

Crown related descriptors were found to contribute most to the PDI (Figure #) with more severely declining trees having reduced crown volume and condition. 
The DAI was a more complex in descriptor contribution with trunk staining and bleeding descriptors contributing highly with bleed prevalence contributing the most. 
Trees with high bleed prevalence and therefore a larger area of their trunk covered with black staining or active bleeds had much higher DAI scores.

@@

>>>>

```{r loadData, include = FALSE}
loadd(site_rf_margin,
      site_rf_post_correction_margin,
      PDI_status_ttest,
      DAI_groups_ttest,
      PDI_rf_tune_params,
      DAI_rf_tune_params,
      analysis_suitable_data,
      site_corrected_analysis_suitable_data,
      remission_PDI_group_ttest,
      remission_DAI_group_ttest)
```

@@

Descriptors were collected relating to the trees shape, size, crown condition and the presence of biotic agents including insects, bacteria and fungi.
All trees were manually assigned a decline status category; control, AOD, COD or remission prior to collection of phenotypic descriptors.
The remission group included trees that had shown stem bleeds associatied with AOD in the past but were not showing signs of stem bleeds at the time of assessment.

Diameter at breast height (DBH) was collected at a stem height of 1.3m, to the nearest 0.1cm.
Total height, timber hight and lowest live crown were measured using an inclinometer to the nearest 0.1m
The timber height was measured at either 7cm top stem diameter or where no main stem could be distinguished.
For the crown, canopy closure, missing crown and absolute crown density percentages were visually estimated within 5% classes

@@

### Oak decline severity indexes of manually assigned decline statuses

A scatter plot of the calculated PDI and DAI scores is shown in Figure # for all 174 trees for which phenotypic descriptor data was collected.
It can clearly be seen that the control trees group towards the lower end of the PDI scale and the sympotmatic COD and AOD groups towards the higher end.
This gave a significant difference when PDI of the control and symptomatic (AOD and COD) decline statuses were compared using an independent two sample t-test (`r glue("t({PDI_status_ttest$parameter}) = {signif(PDI_status_ttest$statistic,digits = 3) %>% abs()}, p < 0.001")`) with group means of `r signif(PDI_status_ttest$estimate1,digits = 3)` and `r signif(PDI_status_ttest$estimate2,digits = 3)` for non-symptomatic and symptomatic statuses respectively.

On the DAI scale, the AOD trees group at the upper end of the scale (> 0) and the COD trees at the lower end (< 0).
This was confirmed with significance found in the DAI between these groups (`r glue("t({DAI_groups_ttest$parameter}) = {signif(DAI_groups_ttest$statistic,digits = 3) %>% abs()}, p < 0.001")`) with group means of `r signif(DAI_groups_ttest$estimate1,digits = 3)` and `r signif(DAI_groups_ttest$estimate2,digits = 3)` for the AOD and COD groups respectively.

The remission trees were found to be significantly different in their PDI scores from both the control (`r glue("t({remission_PDI_group_ttest$parameter[1]}) = {signif(remission_PDI_group_ttest$statistic[1],digits = 3) %>% abs()}, p < 0.001")`) and symptomatic (AOD and COD; `r glue("t({remission_PDI_group_ttest$parameter[2]}) = {signif(remission_PDI_group_ttest$statistic[2],digits = 3) %>% abs()}, p < 0.001")`) groups with means of `r signif(remission_PDI_group_ttest$estimate1[1],digits = 3)`, `r signif(remission_PDI_group_ttest$estimate2[1],digits = 3)` and `r signif(remission_PDI_group_ttest$estimate2[2],digits = 3)` for the control, remission and symptomatic groups respectively.
As a group these trees are showing moderate overall decline.

The remission trees were also found to be significantly different to the AOD (`r glue("t({remission_DAI_group_ttest$parameter[1]}) = {signif(remission_DAI_group_ttest$statistic[1],digits = 3) %>% abs()}, p = < 0.001")`) and COD (`r glue("t({remission_DAI_group_ttest$parameter[2]}) = {signif(remission_DAI_group_ttest$statistic[2],digits = 3) %>% abs()}, p = {round(remission_DAI_group_ttest$p.value[2],3)}")`) groups on the DAI scale with group means of `r signif(remission_DAI_group_ttest$estimate1[1],digits = 3)`, `r signif(remission_DAI_group_ttest$estimate2[1],digits = 3)` and `r signif(remission_DAI_group_ttest$estimate1[2],digits = 3)` for the AOD, remission and COD groups respectively. 
As a group the remission trees are of neutral decline type.

### Overall descriptor contributions and trends to the decline indexes

- Top ranked descriptors contributing to the decline indexes
- Trends of key phenotypic descriptors withing the PDI and DAI measures
- Agrilus exit hole density had no contribution
- Top ranked descriptor index response surfaces

Crown related descriptors were found to contribute most to the PDI (Figure #a) with more severely declining trees having reduced crown volume and condition. 
The DAI was a more complex in descriptor contribution with trunk staining and bleeding descriptors contributing highly with bleed prevalence contributing the most. 
Trees with high bleed prevalence and therefore a larger area of their trunk covered with black staining or active bleeds had much higher DAI scores.

@@

### Localised descriptor contributions and trends within the PDI and DAI models

- Lime results
